### 2018-03-28 linearity-2018/adjoint-models.tex
Imported all of the results for LAMs.
Next I have to add some narative, spruse some of it up, and add proof
summaries.

We first introduce adjoint models, and the two particular adjoint
models that will be used to model the Lambek Calculus with and without
associativity.  However, before we can do this we have to define a
very simple categorical structure known as a magmoidal category with a
unit.
%% \begin{definition}
%%   \label{def:magmoidal-categories}
%%   1
%% \end{definition}

>>>>

In this section we introduce Lambek Adjoint Models (LAMs). Benton's
LNL model consists of a symmetric monoidal adjunction
$F:\cat{C}\dashv\cat{L}:G$ between a cartesian closed category
$\cat{C}$ and a symmetric monoidal closed category $\cat{L}$. LAM
consists of a monoidal adjunction between a symmetric monoidal closed
category and a Lambek category.
\begin{definition}
  \label{def:lambek-category}
  A \textbf{Lambek category} is a monoidal category $(\cat{L},\tri,I',\alpha',\lambda',\rho')$
  with two functors $- \rightharpoonup - : \cat{L}^{\mathsf{op}} \times \cat{L} \mto \cat{L}$ and
  $- \leftharpoonup - : \cat{L} \times \cat{L}^{\mathsf{op}} \mto \cat{L}$ such that the following
  two natural bijections hold:
  \[
  \begin{array}{lllll}
    \Hom{L}{A \tri B}{C} \cong \Hom{L}{A}{B \rightharpoonup C} & \quad &
    \Hom{L}{A \tri B}{C} \cong \Hom{L}{B}{C \leftharpoonup A}\\
  \end{array}
  \]  
\end{definition}

\begin{definition}
  A \textbf{Lambek Adjoint Model (LAM)}, $(\cat{C},\cat{L},F,G,\eta,\varepsilon)$, consists of
  \begin{itemize}
  \item a symmetric monoidal closed category $(\cat{C},\otimes,I,\alpha,\lambda,\rho)$;
  \item a Lambek category $(\cat{L},\tri,I',\alpha',\lambda',\rho')$;
  \item a monoidal adjunction $F:\cat{C}\dashv\cat{L}:G$ with unit $\eta_X:X\rightarrow GFX$ and
        counit $\varepsilon:FG\rightarrow id_\cat{L}$, where $(F:\cat{C}\rightarrow\cat{L}, m)$
        and $(G:\cat{L}\rightarrow\cat{C}, n)$ are monoidal functors.
  \end{itemize}
\end{definition}
\noindent
Following the tradition, we use letters $X$, $Y$, $Z$ for objects in
$\cat{C}$ and $A$, $B$, $C$ for objects in $\cat{L}$. The rest of this
section proves essential properties of any LAM.

\textbf{An isomorphism}. Let $(\cat{C},\cat{L},F,G,\eta,\varepsilon)$
be a LAM, where $(F,\m{})$ and $(G,\n{})$ are monoidal
functors. Similarly as in Benton's LNL model, $\m{X,Y} : FX \tri FY
\mto F(X \otimes Y)$ are components of a natural isomorphism, and
$\m{I} : I' \mto FI$ is an isomorphism. This is essential for deriving certain
rules of CNC logic, such as tensor elimination in natural deduction.

We define the inverses of $\m{X,Y}:FX\tri FY\rightarrow F(X\otimes Y)$
and $\m{I}:I'\rightarrow FI$ as:

\begin{mathpar}
\bfig
  \morphism<1000,0>[\p{X,Y}:F(X\otimes Y)`F(GFX\otimes GFY);F(\eta_X\otimes\eta_Y)]
  \morphism(1000,0)<900,0>[F(GFX\otimes GFY)`FG(FX\tri FY);F\n{FX,FY}]
  \morphism(1900,0)<750,0>[FG(FX\tri FY)`FX\tri FY;\varepsilon_{FX\tri FX}]
\efig
\and
\bfig
  \morphism<500,0>[\p{I}:FI`FGI';F\n{I'}]
  \morphism(500,0)<400,0>[FGI'`I';\varepsilon_{I'}]
\efig
\end{mathpar}

\begin{theorem}
  $\m{X,Y}$ are components of a natural isomorphism and their inverses are $\p{X,Y}$.
\end{theorem}
\begin{proof}
  This proof follows by a diagram chase using the adjunction in the
  definition of a LAM and naturality.
\end{proof}

\textbf{Strong non-commutative monad}. Next we show that the monad on
$\cat{C}$ in LAM is strong but non-commutative. In Benton's LNL model,
the monad on the cartesian closed category is commutative.
\begin{lemma}
\label{lem:monoidal-monad}
The monad induced by any LAM, $GF : \cat{C} \mto \cat{C}$, is monoidal.
\end{lemma}
\noindent
However, the monad is not symmetric becauase the following diagram does not commute, for the
Lambek category $\cat{L}$ is not symmetric.
\begin{mathpar}
\bfig
  \ptriangle/->`->`/<900,400>[
    GFX\otimes GFY`GFY\otimes GFX`G(FX\tri FY);\e{GFX,GFY}`\n{FX,FY}`]
  \morphism(900,400)<900,0>[GFY\otimes GFX`G(FY\tri FX);\n{FY,FX}]
  \dtriangle(900,0)/`->`->/<900,400>[
    G(FY\tri FX)`GF(X\otimes Y)`GF(Y\otimes X);`G\m{Y,X}`GF\e{X,Y}]
  \morphism|b|<900,0>[G(FX\tri FY)`GF(X\otimes Y);G\m{X,Y}]
\efig
\end{mathpar}

\begin{lemma}
  \label{lem:strong-monad}
  The monad on the symmetric monoidal closed category in LAM is strong.
\end{lemma}

The following lemma is adopted from \cite{}.
\begin{lemma}
\label{lem:monad-com-iff-sym}
  Let $\cat{M}$ be a symmetric monoidal category and $T$ be a strong monad on $\cat{M}$. Then
  $T$ is commutative iff it is symmetric.
\end{lemma}

\begin{theorem}
  The monad on the SMCC in LAM is strong but non-commutative.
\end{theorem}
\begin{proof}
  The proof is obvious. Based on Lemma~\ref{lem:strong-monad} and
  Lemma~\ref{lem:monad-com-iff-sym}, the monad is non-commutative. 
\end{proof}

\textbf{Comonad for exchange}.

\begin{lemma}
  The comonad on the Lambek category in a LAM is monoidal.
\end{lemma}

We then show that the co-Eilenberg-Moore category of the comonad $S$ is symmetric monoidal
closed.

\begin{lemma}
  \label{lem:em-exchange}
  Given a LAM $(\cat{C},\cat{L},F,G,\eta,\varepsilon)$ and the comonad $S$ on $\cat{L}$,
  the co-Eilenberg-Moore category $\cat{L}^S$ has an exchange natural transformation
  $\e{A,B}^S:A\tri B\rightarrow B\tri A$.
\end{lemma}

\begin{theorem}
  The co-Eilenberg-Moore category $\cat{L}^S$ of $S$ is symmetric monoidal closed.
\end{theorem}
